On The Impact of Atoms of Confusion in JavaScript Code

### 2021-04-28 1_Introduction.tex
Updates from Overleaf

@@

Source code misunderstandings can occur due to diverse reasons.
For instance, a poorly understood requirement might lead to poorly implemented code.
An intricate algorithm may require extra time to be fully comprehended.
Besides that, programming languages can also be at fault
and specific syntax and semantics of a language can cause misunderstandings. The same applies to some code
idioms and stylistic choices~\cite{DBLP:conf/msr/GopsteinZFC18}.

@@

%This research focuses on existing JavaScript idioms that a previous research work have proved to introduce misunderstanding in \clang.
JavaScript is the most widely used language for web development and the most often adopted language on GitHub~\cite{BugsJS:ICST}. At the same time, JavaScript is infamously known to be ``the worst most popular language in the world''.\footnote{\url{https://github.com/getify/You-Dont-Know-JS/blob/2nd-ed/preface.md}} It is a weakly typed language~\cite{DynamicBehaviorJS:PLDI} that provides object-oriented, functional, and imperative constructs~\cite{JS-CodeSmells:SCAM}, in addition to features such as run-time evaluation~\cite{BugsJS:ICST,ClientSideBugsJS:TSE}, and asynchronous operations. All these 
%diverse characteristics and capabilities
features combined make JavaScript source code fertile ground for errors~\cite{BugsJS:ICST,ClientSideBugsJS:TSE} and confusion~\cite{JSGoodParts:book}.

%In particular, 
Our research aims to help practitioners identify potentially confusing snippets of code and alternative snippets that can make the code clearer. To achieve that, we investigate JavaScript idioms that might negatively impact program understanding. These idioms correspond to a set of candidates for \emph{atoms of confusion} in JavaScript---that is, small
pieces of code that hinder developers' understanding of the source code of a program for which there are functionally equivalent alternatives that are easier
to understand~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}. For instance, consider the pair of code snippets in Listing~\ref{fig:lst01} (adapted from \textsc{NervJS/taro}).
%for presentation purpose only). 
The code on the left-hand side of the figure contains two atoms of confusion discussed in previous work focusing on the \clang
language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: \emph{Logic as Control Flow} and \emph{Post-increment}. On the right-hand side, we present a clean version of the code, that is, without atoms of confusion. In the experiments of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, removing the atoms \emph{Logic as Control Flow} and \emph{Post-increment} improved the accuracy of program understanding tasks in  41\% and 34\%, respectively.

Recent work~\cite{DBLP:journals/ese/MedeirosLAAKRG19,DBLP:conf/sigsoft/GopsteinIYDZYC17,Langhout:2021:ACJ} has explored the impact of atoms of confusion on program understanding, and confirmed they negatively impact source code readability.
However, these previous studies focus exclusively on statically typed languages (e.g., \clang, \cpplang, and Java). However, JavaScript's programming culture differs from that of languages such as \clang and \cpplang, particularly due to the JavaScript's dynamic
capabilities and weaker type system. Thus, we may find that idioms and constructs that are atoms of confusion in those languages are not particularly difficult to understand for JavaScript programmers and vice-versa. Here we investigate a group of atom candidates that stems from the study of Gopstein and colleagues and detail additional JavaScript constructs that might introduce misunderstanding---according to the opinions of JavaScript
developers.

To understand whether an atom of confusion introduces misunderstandings, we surveyed 140 JavaScript developers. In the survey,  their task was to predict the output of short code snippets, with half of them containing atoms of confusion. 
{\color{red}The number of correct answers for the
  \emph{clean version} of the code snippets for 7 out of \na
  atom candidates is at least 15\% higher than the corresponding
  \emph{confusing version}
  of the code snippet (i.e., with an atom candidate). The presence of an atom candidate
  might also affect the time necessary for submitting a correct answer.
  For instance, the presence of the Comma Operator atom increases
  the average time to predict the output by 76.23\%, though the outputs of snippets containing
  the Pre-Increment atom were correctly predicted 38.19\% faster.

  To learn more about the misunderstandings in the snippets used in the survey, we interviewed 15
  experienced professional developers. Participants considered the code snippets without
  the atom candidates easier to understand in 70\% of the cases,
  therefore corroborating the survey findings. The participants also discussed other 
  idioms that might lead developers to misunderstand JavaScript code.
  The occurrence of atoms in open-source repositories might reveal cases where developers
  should be more or less careful when writing code---and open the possibility to implement
  new program transformations for cleaning up confusing code. To identify such cases, we mined popular JavaScript projects
  from GitHub and found four atom candidates that are highly used in JavaScript
  code. The Implicit Predicate atom, for instance, occurs 19.89 times per 1,000 lines of code.}

Altogether, the main contributions of this paper are:
\rb{revisar as contribuicoes quando o artigo estiver mais maduro. por enquanto, vamos ignorar esses bullets.}

\begin{itemize}
    \item A study about the impact of atoms of confusion on a
    dynamic language.
    \item A list of additional atoms of confusion addressing particular 
    JavaScript constructs and idioms.
    \item A library and infrastructure to mine atoms of confusion using 
    code queries (implemented in the .QL language). This library and 
    infrastructure are freely available and might help researchers and 
    practitioners to automate program analysis at a large scale.
\end{itemize}

>>>>

% Source code misunderstandings
% Recent work... For instance, consider the pair...
% However, focus on C, C++ and Java.
% In this paper ==> Javascript!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% INICIO INTRODUCAO - MARCIO
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Source code misunderstandings can occur due to many reasons. For instance, a poorly understood requirement might lead to poorly implemented code.  Besides that, programming languages can also have specific constructs that may lead to misunderstandings. In this context, the same applies to \emph{atoms of confusion}---that is, small pieces of code that hinder developers' understanding of the source code of a program for which there are functionally equivalent alternatives that are easier to understand~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17, DBLP:conf/msr/GopsteinZFC18}.

Recent work~\cite{DBLP:journals/ese/MedeirosLAAKRG19,DBLP:conf/sigsoft/GopsteinIYDZYC17,Langhout:2021:ACJ} has explored the impact of atoms of confusion on program understanding, and confirmed they negatively impact source code readability. For instance, consider the pair of code snippets in Listing~\ref{fig:lst01} (adapted from \textsc{NervJS/taro}). The code at the left-hand side contains two atoms of confusion discussed in previous work focusing on the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: \emph{Logic as Control Flow} and \emph{Post-increment}. At the right-hand side, we present a clean version of the code, that is, without atoms of confusion. According to the experiments of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, removing the atoms \emph{Logic as Control Flow} and \emph{Post-increment} improved the accuracy of program understanding tasks in 41\% and 34\%, respectively.

However, these previous studies focus exclusively on statically typed languages (e.g., \clang, \cpplang, and Java). This way, there is a knowledge gap with respect to the impact of such atoms in dynamic languages such as JavaScript. In this sense, because JavaScript's programming culture differs from that of languages such as \clang and \cpplang, particularly due to the JavaScript's dynamic capabilities and weaker type system, we may find that idioms and constructs that are atoms of confusion in those languages are not particularly difficult to understand for JavaScript programmers and vice-versa. Also, besides being the most widely used language for web development and the most often adopted language on GitHub~\cite{BugsJS:ICST}, JavaScript contains features that combined make the source code fertile ground for errors~\cite{BugsJS:ICST,ClientSideBugsJS:TSE} and confusion~\cite{JSGoodParts:book}, such as weakly typed language~\cite{DynamicBehaviorJS:PLDI}, object-oriented, functional, and imperative constructs~\cite{JS-CodeSmells:SCAM}, and features such as run-time evaluation~\cite{BugsJS:ICST,ClientSideBugsJS:TSE}, and asynchronous operations.

In this way, to better understand the impact of atoms of confusion in JavaScript, in this paper we present a mixed-method research effort based on a survey, on interviews, and on a mining software repositories study. First, we surveyed 140 JavaScript developers. In the survey, their task was to predict the output of short code snippets, with half of them containing atoms of confusion. The number of correct answers for the \emph{clean version} of the code snippets for 7 out of \na atom candidates is at least 15\% higher than the corresponding \emph{confusing version} of the code snippet (i.e., with an atom candidate). The presence of an atom candidate might also affect the time necessary for submitting a correct answer. For instance, the presence of the Comma Operator atom increases the average time to predict the output by 76.23\%, though the outputs of snippets containing the Pre-Increment atom were correctly predicted 38.19\% faster.

To learn more about the misunderstandings in the snippets used in the survey, we also interviewed 15 experienced professional developers. Participants considered the code snippets without the atom candidates easier to understand in 70\% of the cases, therefore corroborating the survey findings. The participants also discussed other idioms that might lead developers to misunderstand JavaScript code. The occurrence of atoms in open-source repositories might reveal cases where developers should be more or less careful when writing code---and open the possibility to implement new program transformations for cleaning up confusing code. To identify such cases, we mined popular JavaScript projects from GitHub and found four atom candidates that are highly used in JavaScript code. The Implicit Predicate atom, for instance, occurs 19.89 times per 1,000 lines of code.

Altogether, the main contributions of this paper are:

\begin{itemize}

    \item A mixed-method research effort---based on a survey, interviews, and mining software repositories---about the impact of atoms of confusion on a
    dynamic language;

    \item A list of additional atoms of confusion addressing particular 
    JavaScript constructs and idioms;

    \item A library and infrastructure to mine atoms of confusion using 
    code queries (implemented in the .QL language). This library and 
    infrastructure are freely available and might help researchers and 
    practitioners to automate program analysis at a large scale.

\end{itemize}

As implications, our results can help researchers and practitioners to better understand atoms of confusion in JavaScript. For example, our findings might alert the JavaScript community to avoid writing code with certain atoms of confusion. Moreover, our mixed study might help researchers and practitioners to design program transformation tools to remove atoms of confusion in JavaScript. They might prioritize atoms that cause a lot of confusion and are very common in the repositories, for instance.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% FINAL INTRODUCAO - MARCIO
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%Here we investigate a group of atom candidates that stems from the study of Gopstein and colleagues and detail additional JavaScript constructs that might introduce misunderstanding---according to the opinions of JavaScript developers.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% ESSE PARAGRAFO ESTAVA NA VERSAO ANTIGA
%Source code misunderstandings can occur due to diverse reasons. For instance, a poorly understood requirement might lead to poorly implemented code. An intricate algorithm may require extra time to be fully comprehended. Besides that, programming languages can also be at fault and specific syntax and semantics of a language can cause misunderstandings. The same applies to some code idioms and stylistic choices~\cite{DBLP:conf/msr/GopsteinZFC18}.

@@

%This research focuses on existing JavaScript idioms that a previous research work have proved to introduce misunderstanding in \clang.

% ESSE PARAGRAFO ESTAVA NA VERSAO ANTIGA
%JavaScript is the most widely used language for web development and the most often adopted language on GitHub~\cite{BugsJS:ICST}. At the same time, JavaScript is infamously known to be ``the worst most popular language in the world''.\footnote{\url{https://github.com/getify/You-Dont-Know-JS/blob/2nd-ed/preface.md}} It is a weakly typed language~\cite{DynamicBehaviorJS:PLDI} that provides object-oriented, functional, and imperative constructs~\cite{JS-CodeSmells:SCAM}, in addition to features such as run-time evaluation~\cite{BugsJS:ICST,ClientSideBugsJS:TSE}, and asynchronous operations. All these features combined make JavaScript source code fertile ground for errors~\cite{BugsJS:ICST,ClientSideBugsJS:TSE} and confusion~\cite{JSGoodParts:book}.

% ESSE PARAGRAFO ESTAVA NA VERSAO ANTIGA
%Our research aims to help practitioners identify potentially confusing snippets of code and alternative snippets that can make the code clearer. To achieve that, we investigate JavaScript idioms that might negatively impact program understanding. These idioms correspond to a set of candidates for \emph{atoms of confusion} in JavaScript---that is, small pieces of code that hinder developers' understanding of the source code of a program for which there are functionally equivalent alternatives that are easier to understand~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}. For instance, consider the pair of code snippets in Listing~\ref{fig:lst01} (adapted from \textsc{NervJS/taro}). The code on the left-hand side of the figure contains two atoms of confusion discussed in previous work focusing on the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: \emph{Logic as Control Flow} and \emph{Post-increment}. On the right-hand side, we present a clean version of the code, that is, without atoms of confusion. In the experiments of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, removing the atoms \emph{Logic as Control Flow} and \emph{Post-increment} improved the accuracy of program understanding tasks in  41\% and 34\%, respectively.

% ESSE PARAGRAFO ESTAVA NA VERSAO ANTIGA
%Recent work~\cite{DBLP:journals/ese/MedeirosLAAKRG19,DBLP:conf/sigsoft/GopsteinIYDZYC17,Langhout:2021:ACJ} has explored the impact of atoms of confusion on program understanding, and confirmed they negatively impact source code readability. However, these previous studies focus exclusively on statically typed languages (e.g., \clang, \cpplang, and Java). However, JavaScript's programming culture differs from that of languages such as \clang and \cpplang, particularly due to the JavaScript's dynamic capabilities and weaker type system. Thus, we may find that idioms and constructs that are atoms of confusion in those languages are not particularly difficult to understand for JavaScript programmers and vice-versa. Here we investigate a group of atom candidates that stems from the study of Gopstein and colleagues and detail additional JavaScript constructs that might introduce misunderstanding---according to the opinions of JavaScript developers.

% \begin{figure*}[thb]
% \noindent\begin{minipage}{.45\textwidth}  
% \begin{lstlisting}[language=JavaScript]
% res = properties.reduce((s, b) => {
%   b && s++
%   return s
% }
% \end{lstlisting}
% \end{minipage}\hfill
% \begin{minipage}{.45\textwidth}
% \begin{lstlisting}[language=JavaScript]
% res = properties.reduce((s, b) => {
%   if(b) {
%     s = s + 1
%   }
%   return s
% }
% \end{lstlisting}
% \end{minipage}
% \captionof{lstlisting}{Code snippet from \textsc{NervJS/taro} project. Line 2 on the left-hand side of the figure contains two
%   atoms of confusion discussed in ~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: \emph{Logic
%   as Control Flow} and \emph{Post-increment}. On the right-hand side, we present an alternative implementation without atoms of confusion. We could simplify the code on the left-hand side even more, by using \texttt{filter} and \texttt{length} instead of the \texttt{reduce} recursive pattern (e.g., \texttt{res = properties.filter(p => p).length}).}\label{fig:lst01}
% \end{figure*}

% ESSE PARAGRAFO ESTAVA NA VERSAO ANTIGA
%To understand whether an atom of confusion introduces misunderstandings, we surveyed 140 JavaScript developers. In the survey,  their task was to predict the output of short code snippets, with half of them containing atoms of confusion. {\color{red}The number of correct answers for the \emph{clean version} of the code snippets for 7 out of \na atom candidates is at least 15\% higher than the corresponding \emph{confusing version} of the code snippet (i.e., with an atom candidate). The presence of an atom candidate might also affect the time necessary for submitting a correct answer. For instance, the presence of the Comma Operator atom increases  the average time to predict the output by 76.23\%, though the outputs of snippets containing the Pre-Increment atom were correctly predicted 38.19\% faster.

% ESSE PARAGRAFO ESTAVA NA VERSAO ANTIGA
%To learn more about the misunderstandings in the snippets used in the survey, we interviewed 15 experienced professional developers. Participants considered the code snippets without the atom candidates easier to understand in 70\% of the cases, therefore corroborating the survey findings. The participants also discussed other idioms that might lead developers to misunderstand JavaScript code. The occurrence of atoms in open-source repositories might reveal cases where developers should be more or less careful when writing code---and open the possibility to implement new program transformations for cleaning up confusing code. To identify such cases, we mined popular JavaScript projects from GitHub and found four atom candidates that are highly used in JavaScript code. The Implicit Predicate atom, for instance, occurs 19.89 times per 1,000 lines of code.

% ESSE PARAGRAFO ESTAVA NA VERSAO ANTIGA
%Altogether, the main contributions of this paper are: \rb{revisar as contribuicoes quando o artigo estiver mais maduro. por enquanto, vamos ignorar esses bullets.}

% \begin{itemize}
%     \item A study about the impact of atoms of confusion on a
%     dynamic language.
%     \item A list of additional atoms of confusion addressing particular 
%     JavaScript constructs and idioms.
%     \item A library and infrastructure to mine atoms of confusion using 
%     code queries (implemented in the .QL language). This library and 
%     infrastructure are freely available and might help researchers and 
%     practitioners to automate program analysis at a large scale.
% \end{itemize}

### 2021-04-28 2_Background.tex
Updates from Overleaf

\section{Background and Related Work}
\label{back}
\subsection{Program Comprehension}

@@

The concept of program comprehension is central to software maintenance and feature development \cite{DBLP:conf/iwpc/TilleySP96}. Regardless of whether they are performing maintenance on legacy software or adding new features, programmers have to understand what the source code does before making any improvements. A natural consequence of this is that a programmer must also have in mind that other programmers (or their future selves) 
%are going to have to
will
spend time trying to understand the code being currently developed before doing any work themselves.

As program comprehension is regarded less as a systematic process than as an objective \cite{DBLP:journals/ibmsj/OHareT94}, there is no single approach or framework that is capable of yielding easily understandable programs. Factors such as problem domain, size and complexity of the code base, and available tools all vary significantly between different settings. These aspects, combined with differences between individuals, which can include experience, working memory capacity, and other cognitive traits, make it clear that program comprehension should be approached in different ways in different contexts.

However, one of the ways in which programs become less comprehensible occurs when programmers favor complicated syntactic constructs instead of using semantically equivalent versions---when these exist---that are potentially less complex. %allowed by the language or framework they are working with. 
This can be a matter of personal taste, since different individuals might hold different opinions on whether a given syntax is difficult to understand. Nevertheless, it is possible to empirically investigate if there exist programming language constructs or idioms that are considerably harder to understand, leading readers of the code to frequently make incorrect predictions of its behavior.

\subsection{Atoms of Confusion}\label{sec:aoc}

The concept of atoms of confusion was introduced by Gopstein et al. \cite{DBLP:conf/msr/GopsteinZFC18}, who defined them as small code patterns that can verifiably lead to misunderstandings. In addition, for a code pattern to be considered an atom, there must be some alternative pattern or language construct that is functionally equivalent and less likely to cause confusion. Their work laid an empirical framework to identify such atoms. In summary, the basic approach to determine whether a code pattern is an atom involves: (i) identifying a set of atom candidates; (ii) identifying alternative constructs to these atom candidates; (iii) building a set of small programs that each contain a single atom candidate; (iv) building different versions of these programs where the atom candidates were removed and which employ the aforementioned alternative constructs; (v) in an experiment or semi-experiment, presenting subjects with the two versions of the programs for each atom candidate and asking subjects to predict their outputs; (vi) statistically comparing the performance of the subjects.
%By having subjects evaluate code where in the expected output relied exclusively on evaluating the result of previously known confusing programs that had been written in \clang or \cpplang, and asking the participants to predict the output of a set of code snippets where the atoms had been removed, the authors were able to isolate and measure the impact of such confusing atoms.

One example that occurs in many programming languages is the Change of Literal Encoding. For instance, the \clang code statement \lstinline{printf("\%d", 013);} 
often leads programmers to predict the output to be \texttt{13}, even though the correct answer is \texttt{11}. This occurs because a leading \texttt{0} in a numerical literal indicates that the number is in base \texttt{8}, a fact that is not only unknown to less experienced programmers, but misleading even for seasoned developers. After formulating a list of 19 atom candidates, Gopstein et al. \cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} were able to identify 15 code idioms and language constructs that present a statistically significant difference in comparative answer correctness when each of the 15 atoms was removed. The least confusing atom showed a 14\% boost in prediction accuracy when the atom was removed, whereas the most confusing one showed a 60\% accuracy increase. Our first study follows the empirical framework discussed in~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, for all atoms that have a counterpart in JavaScript. Gopstein et al. \cite{DBLP:conf/msr/GopsteinZFC18} also performed a qualitative analysis on how confusion manifested when developers inspected code exhibiting different atoms.\castor{I'd rather have two concrete examples here, instead of this.} They note that, for some atoms, a particular incorrect answer was prevalent, whilst there was a variety of incorrect answers for other atoms.

The work of Gopstein et al. \cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} presented the results of comprehensive research on atoms of confusion that can be found in the \clang and \cpplang~languages. Even though JavaScript is a dynamic language that differs from \clang in a variety of aspects, it also has a number of constructs in common. In addition, the imperative aspects of the language are syntactically similar to \clang, e.g., assignments, conditional expressions, \texttt{if}-statements, pre and post increments and decrements, among others. On the one hand, this means that some of the atoms of confusion that exist in \clang programs may also occur in JavaScript code. On the other hand, differences between the languages and the surrounding programming cultures may lead to code patterns that are atoms of confusion in one language not being atoms in the other one.
%Although JavaScript allows us to program in different paradigms from the one C follows, most of the JavaScript code found in popular repositories are written under the imperative paradigm. Also, both languages share many syntactical rules, which allows us to replicate some of the C atoms in JavaScript.

The first study of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} inspired other studies to further investigate the impact of atoms of confusion on code comprehension.
Oliveira et al.~\cite{TheEyesDoNotLie} conducted an experiment using an eye-tracker with 30 participants involving three atoms.
They found that code with atoms of confusion requires more time from participants to predict the output and more visual effort to comprehend.
Gopstein et al., in another paper~\cite{ThinkingAloud} further explored the original atoms by additionally conducting interviews and having programmers discuss among themselves the studied atoms.
The authors argue that this method is important to complement a quantitative analysis.
Indeed, their study revealed findings such as a \textit{``correct evaluation of an atom might not mean that a programmer understood its meaning''
}. We follow a similar approach as we augment our quantitative analysis with interviews.

Gopstein et al. also influenced researchers to look into atoms in languages other than \clang and \cpplang.
Langhout and Aniche~\cite{Langhout:2021:ACJ} derived a set of 14 atoms for Java and performed an experiment with 132 students. They found that participants are more likely to make mistakes (ranging from 2.7 up to 56 times) when atoms are present in half of their studied atoms.
Castor~\cite{castor2018} presented a preliminary catalog of six atom candidates for the Swift programming language. Unlike JavaScript, in Swift, most of the atoms identified by Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} are avoided by construction, e.g., it does not have assignments as values, increment operators, or macros. The paper also presents a structured definition for what it means for a code pattern to be an atom of confusion and a set of general principles to help in the identification of new atoms.

Medeiros et al. \cite{DBLP:journals/ese/MedeirosLAAKRG19} analyzed 50 open source projects written in the C language with the goal of evaluating 12 code patterns called ``misunderstanding patterns'' by the authors. Many of these misunderstanding patterns are either atoms of confusion or atom candidates~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}. This study shows that these patterns are prevalent; among these 50 projects, there are more than 109K occurrences of misunderstanding patterns. 
In order to gauge the relevance of these misunderstanding patterns, the authors sent 35 pull requests removing occurrences of these patterns to randomly-selected open source projects. The authors of the study received feedback for 21 of these pull requests and the maintainers of the projects accepted 8 of them (22.86\%). The authors also analyzed 36 coding style guides from these projects, but found few guidelines specifically addressing misunderstanding patterns.

>>>>

%\section{Background and Related Work}
%\label{back}
%\subsection{Program Comprehension}

@@

% The concept of program comprehension is central to software maintenance and feature development \cite{DBLP:conf/iwpc/TilleySP96}. Regardless of whether they are performing maintenance on legacy software or adding new features, programmers have to understand what the source code does before making any improvements. A natural consequence of this is that a programmer must also have in mind that other programmers (or their future selves) 
% %are going to have to
% will
% spend time trying to understand the code being currently developed before doing any work themselves.

% As program comprehension is regarded less as a systematic process than as an objective \cite{DBLP:journals/ibmsj/OHareT94}, there is no single approach or framework that is capable of yielding easily understandable programs. Factors such as problem domain, size and complexity of the code base, and available tools all vary significantly between different settings. These aspects, combined with differences between individuals, which can include experience, working memory capacity, and other cognitive traits, make it clear that program comprehension should be approached in different ways in different contexts.

% However, one of the ways in which programs become less comprehensible occurs when programmers favor complicated syntactic constructs instead of using semantically equivalent versions---when these exist---that are potentially less complex. %allowed by the language or framework they are working with. 
% This can be a matter of personal taste, since different individuals might hold different opinions on whether a given syntax is difficult to understand. Nevertheless, it is possible to empirically investigate if there exist programming language constructs or idioms that are considerably harder to understand, leading readers of the code to frequently make incorrect predictions of its behavior.

%\subsection{Atoms of Confusion}\label{sec:aoc}

\section{Related Work}

\label{back}

The concept of program comprehension is central to software maintenance and feature development \cite{DBLP:conf/iwpc/TilleySP96, DBLP:journals/ibmsj/OHareT94}. Atoms of confusion is an important field of program comprehension introduced by Gopstein et al. \cite{DBLP:conf/msr/GopsteinZFC18}, who defined them as small code patterns that can verifiably lead to misunderstandings. In addition, for a code pattern to be considered an atom, there must be some alternative pattern or language construct that is functionally equivalent and less likely to cause confusion. One example that occurs in many programming languages is the Change of Literal Encoding. For instance, the \clang code statement \lstinline{printf("\%d", 013);} 
often leads programmers to predict the output to be \texttt{13}, even though the correct answer is \texttt{11}. This occurs because a leading \texttt{0} in a numerical literal indicates that the number is in base \texttt{8}, a fact that is not only unknown to less experienced programmers, but misleading even for seasoned developers. After formulating a list of 19 atom candidates, Gopstein et al. \cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} were able to identify 15 code idioms and language constructs that present a statistically significant difference in comparative answer correctness when each of the 15 atoms was removed. The least confusing atom showed a 14\% boost in prediction accuracy when the atom was removed, whereas the most confusing one showed a 60\% accuracy increase. Our first study follows the empirical framework discussed in~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, for all atoms that have a counterpart in JavaScript. Gopstein et al. \cite{DBLP:conf/msr/GopsteinZFC18} also performed a qualitative analysis on how confusion manifested when developers inspected code exhibiting different atoms.
%\castor{I'd rather have two concrete examples here, instead of this.}
They note that, for some atoms, a particular incorrect answer was prevalent, whilst there was a variety of incorrect answers for other atoms.

%Their work laid an empirical framework to identify such atoms. In summary, the basic approach to determine whether a code pattern is an atom involves: (i) identifying a set of atom candidates; (ii) identifying alternative constructs to these atom candidates; (iii) building a set of small programs that each contain a single atom candidate; (iv) building different versions of these programs where the atom candidates were removed and which employ the aforementioned alternative constructs; (v) in an experiment or semi-experiment, presenting subjects with the two versions of the programs for each atom candidate and asking subjects to predict their outputs; (vi) statistically comparing the performance of the subjects.
%By having subjects evaluate code where in the expected output relied exclusively on evaluating the result of previously known confusing programs that had been written in \clang or \cpplang, and asking the participants to predict the output of a set of code snippets where the atoms had been removed, the authors were able to isolate and measure the impact of such confusing atoms.

Gopstein et al. \cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} also presented the results of a comprehensive research on atoms of confusion that can be found in the \clang and \cpplang~languages. Even though JavaScript is a dynamic language that differs from \clang in a variety of aspects, it also has a number of constructs in common. In addition, the imperative aspects of the language are syntactically similar to \clang, e.g., assignments, conditional expressions, \texttt{if}-statements, pre and post increments and decrements, among others. On the one hand, this means that some of the atoms of confusion that exist in \clang programs may also occur in JavaScript code. On the other hand, differences between the languages and the surrounding programming cultures may lead to code patterns that are atoms of confusion in one language not being atoms in the other one.
%Although JavaScript allows us to program in different paradigms from the one C follows, most of the JavaScript code found in popular repositories are written under the imperative paradigm. Also, both languages share many syntactical rules, which allows us to replicate some of the C atoms in JavaScript.

%The first study of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} inspired other studies to further investigate the impact of atoms of confusion on code comprehension.

Oliveira et al.~\cite{TheEyesDoNotLie} conducted an experiment using an eye-tracker with 30 participants involving three atoms in \clang. They found that code with atoms of confusion requires more time from participants to predict the output and more visual effort to comprehend. Gopstein et al., in another paper~\cite{ThinkingAloud} further explored the original atoms by additionally conducting interviews and having programmers discuss among themselves the studied atoms. The authors argue that this method is important to complement a quantitative analysis. Indeed, their study revealed findings such as a \textit{``correct evaluation of an atom might not mean that a programmer understood its meaning''}. We follow a similar approach as we augment our quantitative analysis with interviews.

%Gopstein et al. also influenced researchers to look into atoms in languages other than \clang and \cpplang.
Langhout and Aniche~\cite{Langhout:2021:ACJ} derived a set of 14 atoms for Java and performed an experiment with 132 students. They found that participants are more likely to make mistakes (ranging from 2.7 up to 56 times) when atoms are present in half of their studied atoms. Castor~\cite{castor2018} presented a preliminary catalog of six atom candidates for the Swift programming language. Unlike JavaScript, in Swift, most of the atoms identified by Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} are avoided by construction, e.g., it does not have assignments as values, increment operators, or macros. 
%The paper also presents a structured definition for what it means for a code pattern to be an atom of confusion and a set of general principles to help in the identification of new atoms.

Medeiros et al. \cite{DBLP:journals/ese/MedeirosLAAKRG19} analyzed 50 open source projects written in the \clang language with the goal of evaluating 12 code patterns called ``misunderstanding patterns'' by the authors. Many of these misunderstanding patterns are either atoms of confusion or atom candidates~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}. This study shows that these patterns are prevalent; among these 50 projects, there are more than 109K occurrences of misunderstanding patterns. 
In order to gauge the relevance of these misunderstanding patterns, the authors sent 35 pull requests removing occurrences of these patterns to randomly-selected open source projects. The authors of the study received feedback for 21 of these pull requests and the maintainers of the projects accepted 8 of them (22.86\%). 
%The authors also analyzed 36 coding style guides from these projects, but found few guidelines specifically addressing misunderstanding patterns.

### 2021-04-28 3_Method.tex
Updates from Overleaf

In the first study, we explore the impact of atom candidates on
understanding JavaScript code (goal of research question RQ1).	
To this end, we investigate whether or not code snippets that
contain atom candidates (a) produce a higher
misunderstanding rate when programmers
try to predict their outcome and (b) require
programmers to take longer to predict their output.
Since JavaScript and C have some constructs in common (Section~\ref{sec:aoc}), we first selected a set of atoms of confusion for the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} that also exist in JavaScript programs. Our supplementary material
summarizes all atoms we consider in our research.

>>>>

In the first study, we explore the impact of atom candidates on
understanding JavaScript code (goal of research question RQ1).	
To this end, we investigate whether or not code snippets that
contain atom candidates (a) produce a higher
misunderstanding rate when programmers
try to predict their outcome and (b) require
programmers to take longer to predict their output.
Since JavaScript and C have some constructs in common (Section~\ref{back}), we first selected a set of atoms of confusion for the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} that also exist in JavaScript programs. Our supplementary material
summarizes all atoms we consider in our research.

### 2021-04-28 6_Conclusion.tex
Updates from Overleaf

\rb{draft. apenas movi esse texto que estava em settings para ca. pode servir de inspiracao para responder algumas das questoes de pesquisa.}

{\color{blue}Based on these results, we provide evidence that, although the atoms of confusion considered in our study do not take a much higher amount of time to predict the output of the code, their impact on program comprehension is, in almost all of our cases, highly significant.}

>>>>

This paper introduced a mixed-method research effort to better understand the impact of atoms of confusion in JavaScript code. First, we conducted a survey with 140 JavaScript developers. We asked them to predict the output of some code snippets (with and without atoms). We provide evidences that, even though the atoms of confusion we considered do not take a much higher amount of time to predict the output of the code, they might still have a significant impact on program comprehension activities in almost all of the cases we studied. Afterwards, we interviewed 15 professional developers. In general, they considered code without atoms of confusion easier to understand. Last but not least, we conducted a mining software repositories study. In particular, we found four atoms of confusion highly used in JavaScript code, i.e., Implicit Predicate, Ternary Operator, Omitted Curly Braces, and Post Increment.

Our efforts have important implications for practice. First, we alert developers to avoid writing JavaScript code with certain atoms of confusion ({\color{red} \textbf{EXEMPLO AQUI???}).} Second, we might help tool developers to create a program transformation tool to remove automatically atoms of confusion. They might firstly focus, for instance, on the atoms we found to be more confusing and at the same time more common in practice.

%\rb{draft. apenas movi esse texto que estava em settings para ca. pode servir de inspiracao para responder algumas das questoes de pesquisa.}

%{\color{blue}Based on these results, we provide evidence that, although the atoms of confusion considered in our study do not take a much higher amount of time to predict the output of the code, their impact on program comprehension is, in almost all of our cases, highly significant.}
